Memory-Efficient Mixed-Precision Implementations\\for Robust Explicit Model Predictive Control

### 2019-06-11 emsoft-2019/V2/conclusion.tex
updates

We have described an automatic technique to tune an explicit MPC
controller that implements the controller in mixed-precision fixed-point arithmetic
while ensuring that the resulting loss of precision does not invalidate the
constraints of the original control problem.
Our technique models potential fixed-point imprecision explicitly as a disturbance
term, and uses robust explicit MPC to design a controller.
A static error analysis and a mixed-precision tuning tool is then 
used to find an efficient implementation of the controller function.
The implementation maintains the numerical error to within
the disturbance bounds used by the robust controller.
In experiments, we show that our technique can yield significant savings in memory,
ranging up to 20\%. 
In addition, the static analysis scales to controllers with hundreds of regions.

>>>>

We have described an automatic technique to tune an explicit MPC
controller that implements the controller in mixed-precision fixed-point arithmetic
while ensuring that the resulting loss of precision does not invalidate the
constraints of the original control problem.
Our technique models potential fixed-point imprecision explicitly as a disturbance
term, and uses robust explicit MPC to design a controller.
A static error analysis and a mixed-precision tuning tool is then 
used to find an efficient implementation of the controller function.
The implementation maintains the numerical error to within
the disturbance bounds used by the robust controller.
In experiments, we show that our technique can yield significant savings in memory,
ranging up to 20\%. 
In addition, the static analysis scales to controllers with hundreds of regions. \Mahmoud{All in all, using our proposed method, model predictive control under input and state constraints can be achieved while the memory footprint is minimized. By leveraging our method, memory requirement for implementation of explicit MPCs over the industrial embedded systems reduces. One exciting direction for future work would be to investigate effect of stochastic noises on the proposed design methodology.}

### 2019-06-11 emsoft-2019/V2/intro.tex
updates

The main difference between MPC and conventional control is in the nature of the function that maps the measured outputs to control actions. 
MPC computes such a function \emph{online}, whereas a conventional controller pre-computes the function offline.
The online computations required in MPC limits its applicability to slow processes and fast computation platforms: 
the sampling time has to be large enough and the platform fast enough to allow enough time for solving the optimization problem
and obtaining the optimal action for the next time instance. 
Moreover, the optimization solver needs to be certified when using MPC in safety critical applications.

One way to tackle these problems is through \emph{explicit MPC} (EMPC) \cite{Bemporad:2002,Alessio2009}, which formulates the optimization
problem but computes offline a symbolic representation of the solution as a function of the state.
At run-time, the solution is evaluated on the current state as in conventional control.
For example, for linear time invariant models with linear constraints and quadratic costs, the optimization
problem for MPC can be modeled as a quadratic program, and EMPC techniques solve the optimization problem
using multi-parametric programming techniques.
The explicit solution is representable as a partition of the controller domain into a number of polyhedral regions,
and an affine map for each region. 
\Mahmoud{Given the current state, EMPC draws the corresponding affine mapping out of a stored lookup table and evaluates the control.}
Explicit MPC thus expands the class of systems being controlled by MPC strategies, by taking out the need to \Mahmoud{online computations}
or to certify a complex optimization routine.

However, there are still bottlenecks in implementing EMPC on resource-constrained embedded micro-controllers.
First, implementations of EMPCs can suffer from large memory usage, because the solution of the optimization
problem can involve many (often hundreds) of regions~\cite{memoryMPC}.
Since the memory consumption grows linearly with the number of regions, 
this can be a limiting factor in using EMPC on resource-constrained micro-controllers.
Second, many low-end micro-controllers only support fixed-point arithmetic.
Errors in the controller implementation are inversely proportional to the number of bits used for representing each variable~\cite{Anta10}. 
An implementation of EMPC has to be \emph{robust} to 
implementation errors to be able to enforce hard constraints on the states at run time~\cite{imperialrmpc}.

@@

Next, we consider the error in the control input arising out of a fixed-point implementation of the solution.
The error has two sources.
First, the fixed point implementation may pick a different polyhedral region (due to imprecision
in checking membership in a polyhedron).
Second, the fixed-precision implementation of the affine function will have a numerical error due to quantization.
We bound both sources of error statically and automatically,  by finding the maximum possible error due to incorrect region selection
as well as a static bound on the error in computing the affine map.
If the error is at most $\Delta$, we know that the implementation satisfies the robustness margin in the model and
we use an automated mixed-precision tuning tool to find an optimal bitwidth allocation that keeps the implementation below the error bound.
If not, we increase the error bound $\Delta$ and run the loop again to find a new robust controller and find its best mixed-precision
implementation.

@@

For smaller examples, the analysis takes only a few minutes of computation.
We also demonstrate the scalability of our mixed-precision tuning and error bounds analysis:
we show that on explicit MPC controllers with thousands of regions, our tool finishes in a few hours.
In absolute terms, the memory saving corresponds to 21KB over a uniform
precision implementation on our largest benchmark.

>>>>

The main difference between MPC and conventional control is in the nature of the function that maps the measured outputs to control actions. 
MPC computes such a function \emph{online}, whereas a conventional controller pre-computes the function offline.
The online computations required in MPC limits its applicability to slow processes and fast computation platforms: 
the sampling time has to be large enough and the platform fast enough to allow enough time for solving the optimization problem
and obtaining the optimal action for the next time instance. 
Moreover, the optimization solver needs to be certified when using MPC in safety critical applications \Mahmoud{\cite{Alessio2009}}.

One way to tackle these problems is through \emph{explicit MPC} (EMPC) \cite{Bemporad:2002,Alessio2009}, which formulates the optimization
problem but computes offline a symbolic representation of the solution as a function of the state.
At run-time, the solution is evaluated on the current state as in conventional control.
For example, for linear time invariant models with linear constraints and quadratic costs, the optimization
problem for MPC can be modeled as a quadratic program, and EMPC techniques solve the optimization problem
using multi-parametric programming techniques.
The explicit solution is representable as a partition of the controller domain into a number of polyhedral regions,
and an affine map for each region. 
\Mahmoud{Given the current state, EMPC draws the corresponding affine mapping out of a stored lookup table and evaluates the control.}
Explicit MPC thus expands the class of systems being controlled by MPC strategies, by taking out the need to \Mahmoud{online computations}
or to certify a complex optimization routine \Mahmoud{\cite{memoryMPC}}.

However, there are still bottlenecks in implementing EMPC on resource-constrained embedded micro-controllers.
First, implementations of EMPCs \Mahmoud{over industrial micro-controllers with limited memory} can suffer from large memory usage, because the solution of the optimization
problem can involve many (often hundreds) of regions~\cite{memoryMPC}.
Since the memory consumption grows linearly with the number of regions, 
this can be a limiting factor in using EMPC on resource-constrained micro-controllers.
Second, many low-end micro-controllers only support fixed-point arithmetic.
Errors in the controller implementation are inversely proportional to the number of bits used for representing each variable~\cite{Anta10}. 
An implementation of EMPC has to be \emph{robust} to 
implementation errors to be able to enforce hard constraints on the states at run time~\cite{imperialrmpc}.

@@

Next, we consider the error in the control input arising out of a fixed-point implementation of the solution.
The error has two sources.
First, the fixed point implementation may pick a different polyhedral region (due to imprecision
in checking membership in a polyhedron).
Second, the fixed-precision implementation of the affine function will have a numerical error due to quantization.
We bound both sources of error statically and automatically,  by finding the maximum possible error due to incorrect region selection
as well as a static bound on the error in computing the affine map.
If the error is at most $\Delta$, we know that the implementation satisfies the robustness margin in the model and
we use an automated mixed-precision tuning tool to find an optimal bitwidth allocation that keeps the implementation below the error bound.
If not, we increase the error bound $\Delta$ and run the loop again to find a new robust controller and find its best mixed-precision
implementation. \Mahmoud{While our formulation is general enough to take the effect of measurement/process noise into account, this is not subject of study of the current paper.}

@@

For smaller examples, the analysis takes only a few minutes of computation.
We also demonstrate the scalability of our mixed-precision tuning and error bounds analysis:
we show that on explicit MPC controllers with thousands of regions, our tool finishes in a few hours.
In absolute terms, the memory saving corresponds to 21KB over a uniform
precision implementation on our largest benchmark. \Mahmoud{It should be noticed that these savings are not affecting the run-time. Synthesis of the controller and the corresponding precisions are all performed offline.}
